The LSST DESC 3x2pt Tomography Optimization Challenge

### 2021-03-12 paper/paper.tex
more text

Figure \ref{fig:funbin_nz} shows the number density obtained by the highest scoring method in
the fiducial metric, {\sc funbins}, for both the riz and griz scenarios.
It is clear from this that fears about restricted photometry catastrophically limiting
the information needed for tomography are unfounded, at least in our idealised scenario.
In both the CosmoDC2 and Buzzard analyses,
even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration 
of overall bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.

@@

For comparison, we also show results using all six LSST bands, ugrizy; these could not be used
with the metacalibration method. The improvement in the sharpness of the bins is noticeable but not large,
and does not translate into large increases in the figures of merit: the DETF score for CosmoDC2
is 169.7, barely higher than the score for griz.  While this is only a general indication and not
a thorough exploration, it is encouraging.

@@

Notably, though, the method that received the highest overall score in Table \ref{tab:cosmodc2}, 
{\sc funbins} did so with a simpler approach - it used a random forest to assign galaxies to bins,
but chose nominal bin edges using an equal width bins in \emph{comoving distance} $\chi$, converting
this to redshift using a fiducial cosmology. The success of this simple solution suggests this
as an easy heuristic for well-optimized nominal edges.

@@

\subsection{ {\sc funbins} }
This method uses the random forest algorithm described in section \ref{sec:randomforest} to assign galaxies to tomographic bins, but modifies the target bin edge selection method. There are 3 options for determining the edges used to assign labels to the training data: \textit{log}, \textit{random}, and \textit{chi}.

@@

\section{Full 9-bin results table}
Tables \ref{tab:full_dc2} and \ref{tab:full_buzz} show all calculated metrics when generating nine tomographic 
bins for CosmoDC2 and Buzzard respectively. The \textit{ww} columns
show weak-lensing metrics, \textit{gg} whow galaxy clustering, and \textit{3x} the full 3x2pt metric.

>>>>

The overall winning method in the 3x2pt case in the fiducial griz 3x2pt CosmoDC2 metric was \textsc{funbins},
which used a random forest to assign bins, but chose nominal bin edges by splitting the range spanned
by the sample into equal co-moving distance bins. As described in \autoref{sec:funbins},  this was designed
to minimze the shot noise for the angular power in each bin, and this has proven successful. 
Figure \ref{fig:funbin_nz} shows the number density obtained by this method, for both the riz and griz 
scenarios, and an additional post-challenge run using the full Rubin ugrizy bands.
It is clear from this that fears about restricted photometry catastrophically limiting
the information needed for tomography are unfounded, at least in our idealised scenario.
In both the CosmoDC2 and Buzzard analyses,
even the $riz$ bands alone are sufficient to split objects into nine reasonably separated
tomographic bins.  Once bins become this narrow the primary analysis concern will become the calibration 
of overall bin $n(z)$ values, rather than initial tomography, so we consider this to be reassuring.

@@

For comparison, we also show results using all six LSST bands, ugrizy; these could not be used
with the metacalibration method. The improvement in the sharpness of the bins is noticeable but not large,
and does not translate into large increases in the figures of merit: the DETF score for CosmoDC2
is 169.7, barely higher than the score for griz, and this is also true for other metrics.   While this is only a general indication and not a thorough exploration, it is encouraging.

\textsc{funbins} seems to be a good and stable general method, though several other methods perform as well
or slightly better in other scenarios.  Notably, many methods reach the same plateau in lensing-only metrics,
suggesting this is somewhat easier to optimize; this is consistent with the general reason that photometric
redshifts suffice for lensing: the broad lensing kernels are less sensitive to small changes.

Several of the 9-bin neural network methods perform well up to seven tomographic bins but fail when run with 
nine.  This is presumably not an intrinsic feature of these methods, but depends on implementation details.
Further development solving these challenges should be valuable, since they are the highest scoring seven-bin
methods.

@@

Notably, though, the winning \textsc{funbins} method used the simpler approach desribed above. The success of this simple and fast solution suggests this as an easy heuristic for well-optimized nominal edges, at least
for cases where clustering-like metrics are important.

@@

\subsection{ {\sc funbins} } \label{sec:funbins}
This method uses the random forest algorithm described in section \ref{sec:randomforest} to assign galaxies to tomographic bins, but modifies the target bin edge selection method. There are 3 options for determining the edges used to assign labels to the training data: \textit{log}, \textit{random}, and \textit{chi}.

@@

\section{Full 9-bin results table}
Tables \ref{tab:full_dc2} and \ref{tab:full_buzz} show all calculated metrics when generating nine tomographic 
bins for CosmoDC2 and Buzzard respectively. The \textit{ww} columns
show weak-lensing metrics, \textit{gg} show galaxy clustering, and \textit{3x} the full 3x2pt metric.
